---
title: "Châtel-Soulet et al. 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo     = TRUE,
                      cache    = TRUE,
                      cache.lazy = FALSE,
                      prompt   = FALSE,
                      tidy     = TRUE,
                      comment  = NA,
                      message  = FALSE,
                      collapse = FALSE)

```


# Environnement

Analysis done within R v4.2.0 and Bioconductor v3.15

```{r library, echo=FALSE}
## All libraries
library(SingleCellExperiment) # v.1.18.0
library(scuttle) # v.1.6.2
library(scDblFinder) # v.1.10.0
library(BiocParallel) # v.1.30.3
library(scran) # v.1.24.0
library(scater) # v.1.24.0
library(batchelor) # v.1.12.3
library(Seurat) # v.4.1.1 
library(SingleR) # v.1.10.0 
library(limma) # v.3.52.2
library(edgeR) # v.3.38.4
library(ggplot2) # v.3.3.6
library(RColorBrewer) # v.1.1-3
```


# Single cell object

With de-multiplexed samples (see Material and method for information about the de-multiplexing)

```{r sce}
## Read RDS object containing single cell object of 55,489 genes and 119,037 cells
sce = readRDS('SingleCellExperiment.rds') # data processed found under GSE261266
```


# Quality Control

QC metrics to identify low-quality cells based on their expression profiles: 

  * The library size
  * The number of expressed features in each cell (mainly genes and mitochondrial genes)
  

```{r QC_metrics, echo=FALSE}
## Basic library properties for each cell
# Mitochondrial genes
is.mito = grepl("^mt-", rowData(sce)$SYMBOL)
# Ribosomal genes
is.ribo = grepl("^Rp(l|s)", rowData(sce)$SYMBOL)
# Cytoplasmic genes
is.cyto =  rowData(sce)$SYMBOL %in% c("Actb", "Tmsb10", "Tmsb4x")

## QC metrics
##  library size, Number of gene expressed, proportion of reads mapped to features
stats = perCellQCMetrics(sce, subsets=list(MT=is.mito, RIBO=is.ribo, CYTO=is.cyto), flatten=FALSE)
sce$scater_qc = stats

## Identifying low-quality cells with adaptive thresholds (see material and method section)

# Library size:
discard.sum = sce$scater_qc$sum < 2511 
# Percentage dicarded:
length(which(discard.sum))/ncol(sce)*100 #  8.35 percent

# Number of gene expressed:
discard.detected = sce$scater_qc$detected < 1250
length(which(discard.detected))/ncol(sce)*100 #  8.1 % discarded

# Cells with more than 0 or max 10% of mitochondrial gene
discard.mito = sce$scater_qc$subsets$MT$percent > 10 | sce$scater_qc$subsets$MT$percent == 0
discard.mito[is.na(discard.mito)] = TRUE
table(discard.mito) # discard 5,955, keep 113,082


# Summary
data.frame(BySum=sum(discard.sum),
           ByFeature=sum(discard.detected),
           ByMito=sum(discard.mito),
           Remaining=sum(!discard.sum & !discard.detected & !discard.mito ))


## Removing low-quality cells
sce = sce[, !discard.sum & !discard.detected & !discard.mito] # 108,503 high quality cells
```

# Doublet detection

```{r doublet_dectection, echo=FALSE}
# Doublet detection by simulation
set.seed(100)
sce = scDblFinder(sce, verbose=TRUE, samples="replicate_id")

# Clean-up metadata
colData(sce)$scDblFinder = colData(sce)[, grepl("scDblFinder", colnames(colData(sce)))]
colData(sce) = colData(sce)[, !grepl("scDblFinder\\.", colnames(colData(sce)))]
```

# Normalisation

Normalization by deconvolution

```{r normalisation, echo=FALSE}
# Change rownames GENEID to SYMBOL
rownames(sce) = rowData(sce)$SYMBOL

# Record average expression
ave.counts = calculateAverage(sce) 
rowData(sce)$ave.count = ave.counts

# Gene high in ambient RNA
to_tag <- vector()
allAmbiant = NULL
for (samp in unique(sce$SampleName)){
   pos = grep(samp, colnames(rowData(sce)))
   ambient = rowData(sce)[,pos]
   ambient = ambient[order(ambient$ambient,decreasing = T),]
   ## Tag everything that has a proportion in the ambient RNA higher than 0.001 (0.1%)
   ambient = ambient[which(ambient$ambient>0.001 & !is.na(ambient$ambient)),]
   ## Tag the genes in sce object
   to_tag <- unique(c(to_tag, row.names(ambient)))
 }
rowData(sce)$high_in_ambient <- rowData(sce)$GENEID %in% to_tag
high_ambiant =  rownames(rowData(sce))[which(rowData(sce)$high_in_ambient==TRUE)]

# Pre-clustering step
bpPar = MulticoreParam(workers = 4, RNGseed = 1234)
set.seed(1234)
clusters = quickCluster(sce, 
                          use.ranks=FALSE, method = "igraph",
                          BSPARAM=BiocSingular::IrlbaParam(),BPPARAM = bpPar)
colData(sce)$quickCluster = factor(clusters) 


# Compute Size factors 
sce = scran::computeSumFactors(sce, 
                                cluster = colData(sce)$quickCluster,
                                subset.row= rowData(sce)$high_in_ambient== FALSE,
                                positive = TRUE,
                                min.mean = 0.1)
# Normalisation
sce = logNormCounts(sce)

```

# Feature selection

```{r feature_selection, echo=FALSE}
# Quantifying technical noise
set.seed(100)
dec.pois = modelGeneVarByPoisson(sce)
## Store it in the single cell object
rowData(sce)$decomposeVar = dec.pois
dec.pois = cbind(dec.pois, data.frame(High_in_ambient=rowData(sce)$high_in_ambient))
## Sort 
dec.pois = dec.pois[order(dec.pois$bio, decreasing=TRUE),]

# Highly variable gene
## Extract HVGs (significant ones with remaining variance > 0.3 & FDR<1%)
hvg.out =  getTopHVGs(dec.pois,fdr.threshold=0.05, var.threshold=0.1) 

# remove the genes, that are high variable and high in the ambient
hvg.out = hvg.out[!hvg.out %in% row.names(sce)[rowData(sce)$high_in_ambient == TRUE]]

# PCA 
set.seed(1000)
sce = runPCA(sce, subset_row=hvg.out)
# Denoise the PCA 
set.seed(1000)
sce = denoisePCA(sce, technical=rowData(sce)$decomposeVar, subset.row=hvg.out)
```



# Correcting batch effects

```{r batch_correction, echo=FALSE}
## MNN correction
set.seed(1234)
mnn.out = fastMNN(sce, batch=sce$batch_full_info, d=50, k=50, subset.row=hvg.out,
                   BSPARAM=BiocSingular::RandomParam(deferred=TRUE))

# runTSNE and UMAP on corrected PCA
bpPar = MulticoreParam(workers = 4, RNGseed = 1234)
mnn.out = runTSNE(mnn.out, dimred="corrected",BPPARAM = bpPar)
mnn.out = runUMAP(mnn.out, dimred="corrected",BPPARAM = bpPar)

## Add corrected PCA/TSNE/UMAP to sce object
reducedDim(sce, "MNN_Sample_k50") <- reducedDim(mnn.out, "corrected")
reducedDim(sce, "TSNE_MNN_Sample_k50") <- reducedDim(mnn.out, "TSNE")
reducedDim(sce, "UMAP_MNN_Sample_k50") <- reducedDim(mnn.out, "UMAP")

```


# Doublet removal

```{r doublet_removal, echo=F}
# remove all cells annotated as doublet (scDblFinder) 
sce = sce[,-which(sce$scDblFinder$scDblFinder.class=='doublet')] # 4508 doublets were removed and 104,013 single cell left. 
```

# Cell cycle assignment

```{r Cyclone, echo=FALSE}
# Using the cyclone() classifier
## reference dataset
mm.pairs <- readRDS(system.file("exdata", "mouse_cycle_markers.rds", package="scran"))
set.seed(100)
assignments <- cyclone(sce, mm.pairs, gene.names=rowData(sce)$GENEID, verbose=T, 
                      BPPARAM=BiocParallel::MulticoreParam(4)) 

# Add info into sce object: 
sce$phases <- assignments$phases
```



# Clustering

main clustering and sub-clustering of cluster 0

```{r clustering, echo=FALSE}
## Main clustering
# convert sce to seurat object
sce.Seurat = sce
sce.Seurat = removeAltExps(sce.Seurat) 
rownames(sce.Seurat) = rowData(sce.Seurat)$GENEID
reducedDims(sce.Seurat) = reducedDims(sce.Seurat)[c('MNN_Sample_k50','TSNE_MNN_Sample_k50','UMAP_MNN_Sample_k50')]
reducedDim(sce.Seurat,'pca') = reducedDim(sce.Seurat,'MNN_Sample_k50') 
reducedDims(sce.Seurat) = reducedDims(sce.Seurat)[c('pca')]
sce.Seurat = as.Seurat(sce.Seurat)

# Find cluster using Seurat functions
set.seed(1234)
# k.param nearest neighbors
sce.Seurat = FindNeighbors(sce.Seurat, dims = 1:20)
# Use knn graph to construct the SNN graph by calculating the neighborhood overlap
sce.Seurat = FindClusters(sce.Seurat, resolution = 0.6)
# add clustering to sce object: 
sce$seurat_res0.6 = Idents(sce.Seurat)


## Sub-clustering cluster 0
# subset the dataset
sce_cluster_2 = sce[,which(sce$seurat_res0.6=='0')]
# Feature selection
set.seed(100)
dec.pois.sce2 = modelGeneVarByPoisson(sce_cluster_2)
## Store it 
rowData(sce_cluster_2)$decomposeVar = dec.pois.sce2
## Sort it
dec.pois.sce2 = dec.pois.sce2[order(dec.pois.sce2$bio, decreasing=TRUE),]
# define HVG for the subset
hvg.out.sce2 = getTopHVGs(dec.pois.sce2,fdr.threshold=0.05, var.threshold=0.1) # first by default= top2000. 
# Rerunning PCA using HVGs only
set.seed(1000)
sce_cluster_2 = runPCA(sce_cluster_2, subset_row=hvg.out.sce2)
# Denoise PCA
set.seed(1000)
sce_cluster_2 = denoisePCA(sce_cluster_2, technical=rowData(sce_cluster_2)$decomposeVar, subset.row=hvg.out.sce2)
# New batch correction
set.seed(1234)
mnn.out = fastMNN(sce_cluster_2, 
                   batch=sce_cluster_2$batch_full_info, d=50, k=50, 
                   subset.row=hvg.out.sce2,
                   BSPARAM=BiocSingular::RandomParam(deferred=TRUE))

# TSNE AND UMAP
bpPar = MulticoreParam(workers = 4, RNGseed = 1234)
mnn.out = runTSNE(mnn.out, dimred="corrected",BPPARAM = bpPar)
mnn.out = runUMAP(mnn.out, dimred="corrected",BPPARAM = bpPar)

# Add to sce object
reducedDim(sce_cluster_2, "MNN_Sample_k50") <- reducedDim(mnn.out, "corrected")
reducedDim(sce_cluster_2, "TSNE_MNN_Sample_k50") <- reducedDim(mnn.out, "TSNE")
reducedDim(sce_cluster_2, "UMAP_MNN_Sample_k50") <- reducedDim(mnn.out, "UMAP")


# Clustering
sce.Seurat.subset = sce_cluster_2
sce.Seurat.subset = removeAltExps(sce.Seurat.subset) 
rownames(sce.Seurat.subset) = rowData(sce.Seurat.subset)$GENEID
reducedDim(sce.Seurat.subset) = reducedDim(sce.Seurat.subset,'MNN_Sample_k50') 
reducedDim(sce.Seurat.subset,'pca') = reducedDim(sce.Seurat.subset,'MNN_Sample_k50') 
reducedDims(sce.Seurat.subset) = reducedDims(sce.Seurat.subset)[c('pca')]
sce.Seurat.subset = as.Seurat(sce.Seurat.subset)

# Find cluster using seurat 
set.seed(1234)
sce.Seurat.subset = FindNeighbors(sce.Seurat.subset, dims = 1:20)
sce.Seurat.subset = FindClusters(sce.Seurat.subset, resolution = 0.3) 
# Seurat cluster identities
sce_cluster_2$clustering_fine_0.3 = as.factor(Idents(sce.Seurat.subset))

## Combine main and sub clusters: 
sce_cluster_2$clustering_fine_0.3 = as.character(sce_cluster_2$clustering_fine_0.3)
sce_cluster_2$clustering_fine_0.3[which(sce_cluster_2$clustering_fine_0.3=='0')] = '0.0'
sce_cluster_2$clustering_fine_0.3[which(sce_cluster_2$clustering_fine_0.3=='1')] = '0.1'
sce_cluster_2$clustering_fine_0.3[which(sce_cluster_2$clustering_fine_0.3=='2')] = '0.2'

## Add final clustering info to sce object
final_clustering = as.character(sce$seurat_res0.6)
names(final_clustering) = colnames(sce)
final_clustering[colnames(sce_cluster_2)] = sce_cluster_2$clustering_fine_0.3
sce$final_clustering = final_clustering
sce$final_clustering = as.factor(sce$final_clustering)
```




# Cell type annotation

Using existing reference

```{r cell_annotation, echo=FALSE}
# Prepare reference dataset (Fast et al. 2021)
mat = fread("exprMatrix.tsv.gz") # We asked the autors for the processed table
meta = read.table("meta.tsv", header=T, sep="\t", as.is=T, row.names=1)
genes = mat[,1][[1]]
genes = gsub(".+[|]", "", genes)
mat = data.frame(mat[,-1], row.names=genes)
colnames(mat) = gsub('.','-',colnames(mat),fixed = T)
# seurat object
so = CreateSeuratObject(counts = mat, project = "myProjectName", meta.data=meta)
sce_fast = as.SingleCellExperiment(so)
table(rownames(sce_fast)%in%rownames(sce))
assays(sce_fast) = assays(sce_fast)['logcounts']

# run singleR on Fast Louvain.Cluste annotation
pred_fast = SingleR(test = sce, ref = sce_fast, labels = sce_fast$Louvain.Cluster)
# add to single cell object
sce$cell_type_fast_Louvain_Cluster = toupper(as.character(pred_fast$pruned.labels)) 
# run singleR on Fast surface marker annotation
pred_fast2 = SingleR(test = sce, ref = sce_fast, labels = sce_fast$surface_marker)
# add to single cell object
sce$cell_type_fast_surface_marker <- toupper(as.character(pred_fast2$pruned.labels)) 
```

# Annotation of clusters

The reasoning behind this annotation can be found in the manuscript 

```{r}
# add the new annotations 
sce$final_annotation = as.character(sce$final_clustering)
sce$final_annotation[sce$final_clustering %in% c("0.0")] = "ST-HSC_1"
sce$final_annotation[sce$final_clustering %in% c("0.2")] = "ST-HSC_2"
sce$final_annotation[sce$final_clustering %in% c("0.1")] = "LT-HSC"
sce$final_annotation[sce$final_clustering %in% c("1")] = "MPP1_1"
sce$final_annotation[sce$final_clustering %in% c("6")] = "MPP1_2"
sce$final_annotation[sce$final_clustering %in% c("2")] = "MPP3_1"
sce$final_annotation[sce$final_clustering %in% c("4")] = "MPP3_2"
sce$final_annotation[sce$final_clustering %in% c("5")] = "MPP3_3"
sce$final_annotation[sce$final_clustering %in% c("8")] = "MPP3_4"
sce$final_annotation[sce$final_clustering %in% c("3")] = "MPP2_1"
sce$final_annotation[sce$final_clustering %in% c("7")] = "MPP2_2"
sce$final_annotation[sce$final_clustering %in% c("11")] = "MPP2_3"
sce$final_annotation[sce$final_clustering %in% c("12")] = "CMP_1"
sce$final_annotation[sce$final_clustering %in% c("16")] = "CMP_2"
sce$final_annotation[sce$final_clustering %in% c("9")] = "GMP_1"
sce$final_annotation[sce$final_clustering %in% c("10")] = "GMP_2"
sce$final_annotation[sce$final_clustering %in% c("13")] = "MEP"
sce$final_annotation[sce$final_clustering %in% c("14")] = "T/NK cells pro"
sce$final_annotation[sce$final_clustering %in% c("15", "19")] = NA
sce$final_annotation[sce$final_clustering %in% c("17")] = "CLP"
sce$final_annotation[sce$final_clustering %in% c("18")] = "B cells pro"

```



# Integrating with protein abundance 

Gating stategy following Fast et al. 2021. See material and method for details

```{r gating}
# ADT expression
adt = altExp(sce, 'ADT')
# log transformation without any normalisation 
logcounts(adt) = log2(counts(adt)+1)
# adt logcount matrix
adt_matrix = data.frame('CD34'=logcounts(adt)['CD34',],'CD150'=logcounts(adt)['CD150',],'CD48'=logcounts(adt)['CD48',])


# Gating of five populations of cells
## +/-: the threshold used are adaptive thresholds and based on the density plot of expression of the ADTs. 

### Thresholds: 
### CD150+: logcount(CD150) > 5.5 & CD150-: logcount(CD150) < 5.5 
### CD48+: logcount(CD48) > 8 & CD150-: logcount(CD48) < 8 

## MPP0: CD150- and CD48- 
gating_fast_MPP0 =  rownames(adt_matrix)[which( adt_matrix[,'CD150'] < 5.5 & adt_matrix[,'CD48'] < 8)] # 10%

## MPP2: CD150+ and CD48+ 
gating_fast_MPP2 =  rownames(adt_matrix)[which( adt_matrix[,'CD150'] > 5.5 & adt_matrix[,'CD48'] > 8)] # 10%

## MPP3/4: CD150- and CD48+ 
gating_fast_MPP34 =  rownames(adt_matrix)[which( adt_matrix[,'CD150'] < 5.5 & adt_matrix[,'CD48'] > 8)] # 43%

## LT-HSC + ST-HSC/MPP1: CD150+ and CD48- 
adt_cd34 = adt_matrix[which( adt_matrix[,'CD150'] > 5.5 & adt_matrix[,'CD48'] < 8),'CD34'] 
### Subdivide this group so that ~80% of them are MPP1 and the rest LT-HSC
cut_30 = sort(adt_cd34,decreasing = F)[round(length(adt_cd34)/100*20)]
### Expression threshold:
negativ = which(adt_cd34 < cut_30)
positiv = which(adt_cd34 >= cut_30)
### LT-HSC
gating_fast_LTHSC =  rownames(adt_matrix)[which( adt_matrix[,'CD34'] < cut_30 & adt_matrix[,'CD150'] > 5.5 & adt_matrix[,'CD48'] < 8)] # 753 (0.72%)
### MPP1
gating_fast_MPP1 = rownames(adt_matrix)[which( adt_matrix[,'CD34'] >= cut_30 & adt_matrix[,'CD150'] > 5.5 & adt_matrix[,'CD48'] < 8)] # 3.3


# Add colData information to sce object
sce$gating_fast = 'no_gate'
sce$gating_fast[which(colnames(sce)%in%gating_fast_LTHSC)] = 'LTHSC'
sce$gating_fast[which(colnames(sce)%in%gating_fast_MPP1)] = 'STHSC'
sce$gating_fast[which(colnames(sce)%in%gating_fast_MPP2)] = 'MPP2'
sce$gating_fast[which(colnames(sce)%in%gating_fast_MPP34)] = 'MPP3'
sce$gating_fast[which(colnames(sce)%in%gating_fast_MPP0)] = 'MPP0'


# color accorinding to annotation
color_vector = sce$gating_fast
color_vector[which(color_vector=='MPP2')] = "#4363d8" 
color_vector[which(color_vector=='MPP3')] = "#3cb44b" 
color_vector[which(color_vector=='no_gate')] = "grey90" 
color_vector[which(color_vector=='LTHSC')] = "#e6194b"
color_vector[which(color_vector=='STHSC')] = "#f58231"
color_vector[which(color_vector=='MPP0')] = "#e5f531"

```


# Cluster filtering

Cluster 17, 18 and 19 were removed from the dataset prior to differential expression analysis (see material and method)

```{r filtering, echo=FALSE}
sce = sce[,-which(sce$final_clustering%in%c('17','18','19'))]
```

# Differential analysis

Pseudo-bulk analysis for each the cluster

```{r Differential_expression, echo=FALSE}
# Pseudo-replicate for pseudo-bulk analysis
sce$full_info_rev = paste0(sce$day,'_',sce$condition,'_',sce$batchNr)
# Rename some cluster (alphabetic sorting)
sce$final_clustering = paste('clust',sce$final_clustering ,sep='')
sce$final_clustering[which(sce$final_clustering=='clust0.0')] = 'clust000'
sce$final_clustering[which(sce$final_clustering=='clust0.1')] = 'clust001'
sce$final_clustering[which(sce$final_clustering=='clust0.2')] = 'clust002'
sce$final_clustering[which(sce$final_clustering=='clust1')] = 'clust01'
sce$final_clustering[which(sce$final_clustering=='clust2')] = 'clust02'
sce$final_clustering[which(sce$final_clustering=='clust3')] = 'clust03'
sce$final_clustering[which(sce$final_clustering=='clust4')] = 'clust04'
sce$final_clustering[which(sce$final_clustering=='clust5')] = 'clust05'
sce$final_clustering[which(sce$final_clustering=='clust6')] = 'clust06'
sce$final_clustering[which(sce$final_clustering=='clust7')] = 'clust07'
sce$final_clustering[which(sce$final_clustering=='clust8')] = 'clust08'
sce$final_clustering[which(sce$final_clustering=='clust9')] = 'clust09'
sce$final_clustering = as.factor(sce$final_clustering)

# Split the data according to cluster+pseudo-replicate
splitCells = factor(paste0(make.names(sce$final_clustering), ".", sce$full_info_rev))
num_cells = table(splitCells)

# Filter: Exclude all samples (cluster+pseudo-replicate) with less than 20 cell (minimum threshold for pseudo-bulk)
excludedCluster = names(num_cells[num_cells < 20]) 
# Remove all cell belonging to these samples 
sce.cell = sce[,-which(splitCells%in%excludedCluster)]
splitCells = factor(paste0(make.names(sce.cell$final_clustering), ".", sce.cell$full_info_rev))

# number of cells per samples and condition
num_cells = table(splitCells)

## Sum of all cell from each sample
sum_by_cluster = vapply(split(colnames(sce.cell), splitCells), function(x){ Matrix::rowSums(counts(sce.cell)[, x]) }, numeric(length = nrow(counts(sce.cell)))) 

## Metadata info: Create data frame containing information about the samples 
pheno = data.frame(summed_sample=colnames(sum_by_cluster))
row.names(pheno = pheno$summed_sample)
pheno$num_cells = num_cells
pheno.comp = pheno
# cluster + day
clu.day = unlist(lapply(pheno$summed_sample, function(x) strsplit(x,'_')[[1]][1]))
# cluster 
pheno.comp$clu = unlist(lapply(pheno$summed_sample, function(x) strsplit(x,'.',fixed = T)[[1]][1]))
# sample name
pheno.comp$sam = unlist(lapply(pheno$summed_sample, function(x) strsplit(x,'.',fixed = T)[[1]][2]))
# day 
pheno.comp$day = unlist(lapply(pheno.comp$sam, function(x) strsplit(x,'_',fixed=T)[[1]][1]))
# treatment
pheno.comp$tpo = unlist(lapply(pheno.comp$sam, function(x) strsplit(x,'_',fixed=T)[[1]][2]))
# fusion/no fusion
pheno.comp$fusion =unlist(lapply(pheno.comp$sam, function(x) strsplit(x,'_',fixed=T)[[1]][3]))
# full condition: day + treatment + fusion
pheno.comp$condition = paste(pheno.comp$day,pheno.comp$tpo,pheno.comp$fusion,sep='_')
# Add these info to final data frame
pheno$cluster = factor(pheno.comp$clu)
pheno$replicate = factor(pheno.comp$sam)
pheno$condition = factor(pheno.comp$condition)
```

```{r DGEList, echo=FALSE}
## Create DGEList object 
# Create unique row names
row.names(sum_by_cluster) = uniquifyFeatureNames(rowData(sce)$GENEID, rowData(sce)$SYMBOL)
# combine info into dgelist
dge = DGEList(counts=sum_by_cluster,
             group=factor(paste(pheno$cluster, pheno$condition, sep=".")),
             samples=pheno,
             genes=as.data.frame(rowData(sce.cell)[, c('GENEID','SYMBOL','ENTREZID','uniqueSymbol','DESCRIPTION')]), 
             remove.zeros=FALSE)

# calculate normalization factors
dge = calcNormFactors(dge)

## Gene filtering
# Keep only the genes that are expressed in at least 2 samples
keep = rowSums(edgeR::cpm(dge) > 1) >= 2
# Subset, while updating total counts
dge = dge[keep, , keep.lib.sizes=FALSE] 
# Redo normalization (TMM)
dge = calcNormFactors(dge) 
# Clean-up 
row.names(dge) = dge$genes$GENEID 
# keep only most informative information
dge$genes = dge$genes[, c(1, 4, 3, 5)]
names(dge$genes) = c("GENEID", "SYMBOL", "ENTREZID", "DESCRIPTION")
# Visualize CPM after filtering
log2cpm.raw = edgeR::cpm(dge, log=TRUE, prior.count=8, normalized.lib.sizes=FALSE)
# and after TMM normalization
log2cpm = edgeR::cpm(dge, log=TRUE, prior.count=8, normalized.lib.sizes=TRUE) 


## design matrix 
moma = model.matrix(~ 0 + group, data=dge$samples) 
colnames(moma) = gsub("group", "", colnames(moma)) 
colnames(moma) = make.names(colnames(moma))


## Define cluster names :
clusters_day2 <- c(
  sprintf("clust%03d", 0:2),  # clust000, clust001, clust002
  sprintf("clust%02d", 1:16)  # clust01 to clust14
)

##cluster 0.02 has not enough replicate at day5:
clusters_day5 <- setdiff(clusters_day2, "clust002")


## Define experimental conditions for contrast
conditions <- c("TPO_FUSION", "PBS_FUSION")

## Function to build contrast definitions per day
make_contrasts_per_day <- function(day, clusters) {
  expand.grid(cluster = clusters, stringsAsFactors = FALSE) |>
    dplyr::mutate(
      contrast_name = paste0(cluster, ".", day),
      contrast_formula = paste0(cluster, ".", day, "_", conditions[1],
                                " - ", cluster, ".", day, "_", conditions[2])
    )
}

## Generate contrasts for each day
contrasts_day2 <- make_contrasts_per_day("day2", clusters_day2)
contrasts_day5 <- make_contrasts_per_day("day5", clusters_day5)

## days to generate
days <- list(day2 = clusters_day2, day5 = clusters_day5)

# helper to generate the four contrasts per cluster/day with exact naming
build_contrasts_for <- function(cluster, day) {
  base <- paste0(cluster, ".", day)        # e.g. "clust000.day2"
  names_vec <- c(
    paste0(cluster, ".fusion.", day),    # e.g. "clust000.fusion.day2"
    paste0(cluster, ".nofusion.", day),
    paste0(cluster, ".tpo.", day),
    paste0(cluster, ".pbs.", day)
  )
  formulas <- c(
    paste0(base, "_TPO_FUSION - ", base, "_PBS_FUSION"),         # fusion: TPO_FUSION - PBS_FUSION
    paste0(base, "_TPO_noFUSION - ", base, "_PBS_noFUSION"),     # nofusion: TPO_noFUSION - PBS_noFUSION
    paste0(base, "_TPO_FUSION - ", base, "_TPO_noFUSION"),       # tpo: TPO_FUSION - TPO_noFUSION
    paste0(base, "_PBS_FUSION - ", base, "_PBS_noFUSION")        # pbs: PBS_FUSION - PBS_noFUSION
  )
  setNames(formulas, names_vec)
}

## Build the full named vector of contrasts
contrast.list <- unlist(lapply(names(days), function(d) {
  clusts <- days[[d]]
  unlist(lapply(clusts, function(cl) build_contrasts_for(cl, d)))
}), use.names = TRUE)



## build the contrast matrix
contrasts.matrix <- makeContrasts(contrasts = contrast.list, levels = moma)
colnames(contrasts.matrix) = names(contrast.list)


## Differential expression per cluster
# save the results per cluster in fit list
fits = list()

# Loop across the contrast
for (i in 1:ncol(contrasts.matrix)) {
  
  # Name of the constrast
  contr.name = colnames(contrasts.matrix)[i]
  
  # Print info
  message("   Working on ... ", contr.name)
  
  # Subset samples belonging to the contrast
  dge.temp = dge[, dge$samples$group %in% names(which(contrasts.matrix[,i] != 0))]
  
  # Subset genes expressed in at least 2 replicates 
  keep = rowSums(edgeR::cpm(dge.temp) > 1) >= 2  
  # Redo normalization 
  dge.temp = dge.temp[keep, , keep.lib.sizes=FALSE] 
  
  # Filtering step on the minimum number of cells with expression
  sce.temp = 
    sce[
      rowData(sce)$GENEID %in% dge.temp$genes$GENEID, 
      paste0(sce$final_clustering, ".",sce$day,'_', sce$TPO,'_',sce$Fusion) %in% names(which(contrasts.matrix[,i] != 0)) ]
  
  ## Should be detected in at least 5% of the cells
  keep = rowSums(counts(sce.temp) > 0 ) > dim(sce.temp)[2]/100*5
  dge.temp = dge.temp[keep, , keep.lib.sizes=FALSE] 
  

  # Redo normalization 
  dge.temp = calcNormFactors(dge.temp) 
  
  # logExpression
  log2cpm.temp = edgeR::cpm(dge.temp, log=TRUE, prior.count=8, normalized.lib.sizes=TRUE)

  # Design matrix 
  moma.temp = model.matrix(~ 0 + group, data=dge.temp$samples) 
  # Rename column (readability)
  colnames(moma.temp) = gsub("group", "", colnames(moma.temp)) 
  colnames(moma.temp) = make.names(colnames(moma.temp))

  
  # Estimate dispersion
  dge.temp = estimateDisp(dge.temp, moma.temp)
  
  # Fit a negative binomial generalized log-linear model
  fit1 = glmFit(dge.temp, moma.temp, prior.count=8) 
  lrt = glmLRT(fit1, contrast=contrasts.matrix[names(which(contrasts.matrix[,i] != 0)), i])

  ## Add contrast's results to list
  fits[[contr.name]] = lrt
  rm(lrt)
  fits[[contr.name]]$dge = dge.temp
  fits[[contr.name]]$moma = moma.temp
}


```

# Enrichement analysis

CAMERA (within Limma framework) statistics using MSigDB database v7.5.1


```{r GSEA, echo=FALSE}
## Sets can be download here https://www.gsea-msigdb.org/gsea/msigdb (!!MSigDB v7.5.1)
## load the downloaded sets within R 
ff <- list.files(gseaDir, pattern="\\.CATEGORY.*ensembl.mmu\\.GeneIdsList\\.rds")
for (f in ff) {
    assign(sub("\\.GeneIdsList\\.rds","",f), readRDS(sprintf("%s/%s", gseaDir, f)))
}


# store CAMERA results
res <- list()

# Loop across catergory
for (f in sub("\\.GeneIdsList\\.rds","",ff)) {

  res[[f]] <- list()
  message("Working on ... ", f)
  for (i in 1:ncol(contrasts.matrix)) {
    contr.name <- colnames(contrasts.matrix)[i]
    message("   Working on ... ", contr.name)
    
    ## process and filter msigDB gene sets
    indx <- ids2indices(gene.sets=get(f), identifiers=fits[[contr.name]]$dge$genes$GENEID)
    indx <- indx[sapply(indx, length) >= 5] 
    
    ## add CAMERA stats, using MsigDb
    ## FDR: Benjamini and Hochberg FDR adjusted p-value.
    res[[f]][[contr.name]] = camera(y = fits[[contr.name]]$dge,
              index=indx, 
              design = fits[[contr.name]]$moma, 
              inter.gene.cor = 0.01,
              contrast=fits[[contr.name]]$contrasts,
              trend.var=FALSE,
              weights=NULL, 
              use.ranks = FALSE,
              fit=NULL,
              allow.neg.cor=FALSE,
              sort=TRUE)
    
  
    }
  
}
    






```

# Reproducibility


```{r}


sessionInfo()

```

