---
title: "Châtel-Soulet et al. 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo     = TRUE,
                      cache    = TRUE,
                      cache.lazy = FALSE,
                      prompt   = FALSE,
                      tidy     = TRUE,
                      comment  = NA,
                      message  = FALSE,
                      collapse = FALSE)

```


# Environnement

```{r library, echo=FALSE}
library(beepr)
library(RColorBrewer)
library(viridis)
library(gplots)
library(pheatmap)
library(ComplexHeatmap)
library(circlize)
library(ggplot2)
library(ggrepel)
library(tidyverse)
library(org.Mm.eg.db)
library(ensembldb)
library(edgeR)
library(limma)
library(SingleCellExperiment)
library(scater)
library(scran)
library(mvoutlier)
library(DropletUtils)
library(SingleR)
library(Seurat)
library(dynamicTreeCut)
library(WGCNA)
library(cluster)
library(batchelor)
library(RSQLite)
library(scDblFinder)
library(randomcoloR)
library(batchelor)
library(scuttle)
library(nortest)
library(ggpubr)
library(bluster)
library(vioplot)
library(GEOquery)
library(moe430a.db)
library(BiocParallel)
library(AUCell)
library(GSEABase)
```


# Single cell object

With de-multiplexed samples (see Material and method for information about the de-multiplexing)

```{r sce}
## Read RDS object containing single cell object
sce = readRDS('SingleCellExperiment.rds') # data processed found under GSEXXXXXX
```


# Quality Control

QC metrics to identify low-quality cells based on their expression profiles: 

  * The library size
  * The number of expressed features in each cell (mainly genes and mitochondrial genes)
  

```{r QC_metrics, echo=FALSE}
## Basic library properties for each cell
# Mitochondrial genes
is.mito = grepl("^mt-", rowData(sce)$SYMBOL)
# Ribosomal genes
is.ribo = grepl("^Rp(l|s)", rowData(sce)$SYMBOL)
# Cytoplasmic genes
is.cyto =  rowData(sce)$SYMBOL %in% c("Actb", "Tmsb10", "Tmsb4x")

## QC metrics
##  library size, Number of gene expressed, proportion of reads mapped to features
stats = perCellQCMetrics(sce, subsets=list(MT=is.mito, RIBO=is.ribo, CYTO=is.cyto), flatten=FALSE)
sce$scater_qc = stats

## Identifying low-quality cells with adaptive thresholds (see material and method section)

# Library size:
discard.sum = sce$scater_qc$sum < 2511 
# Percentage dicarded:
length(which(discard.sum))/ncol(sce)*100 #  8.35 percent

# Number of gene expressed:
discard.detected = sce$scater_qc$detected < 1250
length(which(discard.detected))/ncol(sce)*100 #  8.1 % discarded

# Cells with more than 0 or max 10% of mitochondrial gene
discard.mito = sce$scater_qc$subsets$MT$percent > 10 | sce$scater_qc$subsets$MT$percent == 0
discard.mito[is.na(discard.mito)] = TRUE
table(discard.mito) # discard 5,955, keep 113,082


# Summary
data.frame(BySum=sum(discard.sum),
           ByFeature=sum(discard.detected),
           ByMito=sum(discard.mito),
           Remaining=sum(!discard.sum & !discard.detected & !discard.mito ))


## Removing low-quality cells
sce = sce[, !discard.sum & !discard.detected & !discard.mito]
```

# Doublet detection

```{r doublet_dectection, echo=FALSE}
# Doublet detection by simulation
set.seed(100)
sce = scDblFinder(sce, verbose=TRUE, samples="replicate_id")

# Clean-up metadata
colData(sce)$scDblFinder = colData(sce)[, grepl("scDblFinder", colnames(colData(sce)))]
colData(sce) = colData(sce)[, !grepl("scDblFinder\\.", colnames(colData(sce)))]
```

# Normalisation

Normalization by deconvolution

```{r normalisation, echo=FALSE}
# Change rownames GENEID to SYMBOL
rownames(sce) = rowData(sce)$SYMBOL

# Record average expression
ave.counts = calculateAverage(sce) 
rowData(sce)$ave.count = ave.counts

# Gene high in ambient RNA
to_tag <- vector()
allAmbiant = NULL
for (samp in unique(sce$SampleName)){
   pos = grep(samp, colnames(rowData(sce)))
   ambient = rowData(sce)[,pos]
   ambient = ambient[order(ambient$ambient,decreasing = T),]
   ## Tag everything that has a proportion in the ambient RNA higher than 0.001 (0.1%)
   ambient = ambient[which(ambient$ambient>0.001 & !is.na(ambient$ambient)),]
   ## Tag the genes in sce object
   to_tag <- unique(c(to_tag, row.names(ambient)))
 }
rowData(sce)$high_in_ambient <- rowData(sce)$GENEID %in% to_tag
high_ambiant =  rownames(rowData(sce))[which(rowData(sce)$high_in_ambient==TRUE)]

# Pre-clustering step
bpPar = MulticoreParam(workers = 4, RNGseed = 1234)
set.seed(1234)
clusters = quickCluster(sce, 
                          use.ranks=FALSE, method = "igraph",
                          BSPARAM=BiocSingular::IrlbaParam(),BPPARAM = bpPar)
colData(sce)$quickCluster = factor(clusters) 


# Compute Size factors 
sce = scran::computeSumFactors(sce, 
                                cluster = colData(sce)$quickCluster,
                                subset.row= rowData(sce)$high_in_ambient== FALSE,
                                positive = TRUE,
                                min.mean = 0.1)
# Normalisation
sce = logNormCounts(sce)

```

# Feature selection

```{r feature_selection, echo=FALSE}
# Quantifying technical noise
set.seed(100)
dec.pois = modelGeneVarByPoisson(sce)
## Store it in the single cell object
rowData(sce)$decomposeVar = dec.pois
dec.pois = cbind(dec.pois, data.frame(High_in_ambient=rowData(sce)$high_in_ambient))
## Sort 
dec.pois = dec.pois[order(dec.pois$bio, decreasing=TRUE),]

# Highly variable gene
## Extract HVGs (significant ones with remaining variance > 0.3 & FDR<1%)
hvg.out =  getTopHVGs(dec.pois,fdr.threshold=0.05, var.threshold=0.1) 

# remove the genes, that are high variable and high in the ambient
hvg.out = hvg.out[!hvg.out %in% row.names(sce)[rowData(sce)$high_in_ambient == TRUE]]

# PCA 
set.seed(1000)
sce = runPCA(sce, subset_row=hvg.out)
# Denoise the PCA 
set.seed(1000)
sce = denoisePCA(sce, technical=rowData(sce)$decomposeVar, subset.row=hvg.out)

```



# Correcting batch effects

```{r batch_correction, echo=FALSE}
## MNN correction
set.seed(1234)
mnn.out = fastMNN(sce, batch=sce$batch_full_info, d=50, k=50, subset.row=hvg.out,
                   BSPARAM=BiocSingular::RandomParam(deferred=TRUE))

# runTSNE and UMAP on corrected PCA
bpPar = MulticoreParam(workers = 4, RNGseed = 1234)
mnn.out = runTSNE(mnn.out, dimred="corrected",BPPARAM = bpPar)
mnn.out = runUMAP(mnn.out, dimred="corrected",BPPARAM = bpPar)

## Add corrected PCA/TSNE/UMAP to sce object
reducedDim(sce, "MNN_Sample_k50") <- reducedDim(mnn.out, "corrected")
reducedDim(sce, "TSNE_MNN_Sample_k50") <- reducedDim(mnn.out, "TSNE")
reducedDim(sce, "UMAP_MNN_Sample_k50") <- reducedDim(mnn.out, "UMAP")

```


# Doublet removal

```{r doublet_removal, echo=F}
# remove all cells annotated as doublet (scDblFinder) 
sce = sce[,-which(sce$scDblFinder$scDblFinder.class=='doublet')] # 4508 doublet
```

# Cell cycle assignment

```{r Cyclone, echo=FALSE}
# Using the cyclone() classifier
## reference dataset
mm.pairs <- readRDS(system.file("exdata", "mouse_cycle_markers.rds", package="scran"))
set.seed(100)
assignments <- cyclone(sce, mm.pairs, gene.names=rowData(sce)$GENEID, verbose=T, 
                      BPPARAM=BiocParallel::MulticoreParam(4)) 

# Add info into sce object: 
sce$phases <- assignments$phases
```



# Clustering

main clustering and sub-clustering of cluster 0

```{r clustering, echo=FALSE}
## Main clustering
# convert sce to seurat object
sce.Seurat = sce
sce.Seurat = removeAltExps(sce.Seurat) 
rownames(sce.Seurat) = rowData(sce.Seurat)$GENEID
reducedDims(sce.Seurat) = reducedDims(sce.Seurat)[c('MNN_Sample_k50','TSNE_MNN_Sample_k50','UMAP_MNN_Sample_k50')]
reducedDim(sce.Seurat,'pca') = reducedDim(sce.Seurat,'MNN_Sample_k50') 
reducedDims(sce.Seurat) = reducedDims(sce.Seurat)[c('pca')]
sce.Seurat = as.Seurat(sce.Seurat)

# Find cluster using seurat
set.seed(1234)
# k.param nearest neighbors
sce.Seurat = FindNeighbors(sce.Seurat, dims = 1:20)
# Use knn graph to construct the SNN graph by calculating the neighborhood overlap
sce.Seurat = FindClusters(sce.Seurat, resolution = 0.6)
# add clustering to sce object: 
sce$seurat_res0.6 = Idents(sce.Seurat)


## Sub-clustering cluster 0
# subset the dataset
sce_cluster_2 = sce[,which(sce$seurat_res0.6=='0')]
# Feature selection
set.seed(100)
dec.pois.sce2 = modelGeneVarByPoisson(sce_cluster_2)
## Store it 
rowData(sce_cluster_2)$decomposeVar = dec.pois.sce2
## Sort it
dec.pois.sce2 = dec.pois.sce2[order(dec.pois.sce2$bio, decreasing=TRUE),]
# define HVG for the subset
hvg.out.sce2 = getTopHVGs(dec.pois.sce2,fdr.threshold=0.05, var.threshold=0.1) # first by default= top2000. 
# Rerunning PCA using HVGs only
set.seed(1000)
sce_cluster_2 = runPCA(sce_cluster_2, subset_row=hvg.out.sce2)
# Denoise PCA
set.seed(1000)
sce_cluster_2 = denoisePCA(sce_cluster_2, technical=rowData(sce_cluster_2)$decomposeVar, subset.row=hvg.out.sce2)
# New batch correction
set.seed(1234)
mnn.out <- fastMNN(sce_cluster_2, 
                   batch=sce_cluster_2$batch_full_info, d=50, k=50, 
                   subset.row=hvg.out.sce2,
                   BSPARAM=BiocSingular::RandomParam(deferred=TRUE))

# TSNE AND UMAP
bpPar = MulticoreParam(workers = 4, RNGseed = 1234)
mnn.out = runTSNE(mnn.out, dimred="corrected",BPPARAM = bpPar)
mnn.out = runUMAP(mnn.out, dimred="corrected",BPPARAM = bpPar)
# Add to sce object
reducedDim(sce_cluster_2, "MNN_Sample_k50") <- reducedDim(mnn.out, "corrected")
reducedDim(sce_cluster_2, "TSNE_MNN_Sample_k50") <- reducedDim(mnn.out, "TSNE")
reducedDim(sce_cluster_2, "UMAP_MNN_Sample_k50") <- reducedDim(mnn.out, "UMAP")


# Clustering
sce.Seurat.subset = sce_cluster_2
sce.Seurat.subset = removeAltExps(sce.Seurat.subset) 
rownames(sce.Seurat.subset) = rowData(sce.Seurat.subset)$GENEID
reducedDim(sce.Seurat.subset) = reducedDim(sce.Seurat.subset,'MNN_Sample_k50') 
sce.Seurat.subset = as.Seurat(sce.Seurat.subset)

# Find cluster using seurat 
set.seed(1234)
sce.Seurat.subset <- FindNeighbors(sce.Seurat.subset, dims = 1:20)
sce.Seurat.subset <- FindClusters(sce.Seurat.subset, resolution = 0.3) 
# Seurat cluster identities
sce_cluster_2$clustering_fine_0.3 = as.factor(Idents(sce.Seurat.subset))

## Combine main and sub clusters: 
sce_cluster_2$clustering_fine_0.3 = as.character(sce_cluster_2$clustering_fine_0.3)
sce_cluster_2$clustering_fine_0.3[which(sce_cluster_2$clustering_fine_0.3=='0')] = '0.0'
sce_cluster_2$clustering_fine_0.3[which(sce_cluster_2$clustering_fine_0.3=='1')] = '0.1'
sce_cluster_2$clustering_fine_0.3[which(sce_cluster_2$clustering_fine_0.3=='2')] = '0.2'

combined_clustering = as.character(sce$seurat_res0.6)
names(combined_clustering) = colnames(sce)
combined_clustering[colnames(sce_cluster_2)] = sce_cluster_2$clustering_fine_0.3
sce$combined_clustering = combined_clustering
sce$combined_clustering = as.factor(sce$combined_clustering)
```




# Cell type annotation

Using existing references

```{r cell_annotation, echo=FALSE}
# Prepare reference dataset (Fast et al. 2021)
mat = fread("exprMatrix.tsv.gz") # We asked the autors for the processed table
meta = read.table("meta.tsv", header=T, sep="\t", as.is=T, row.names=1)
genes = mat[,1][[1]]
genes = gsub(".+[|]", "", genes)
mat = data.frame(mat[,-1], row.names=genes)
colnames(mat) = gsub('.','-',colnames(mat),fixed = T)
so = CreateSeuratObject(counts = mat, project = "myProjectName", meta.data=meta)
sce_fast = as.SingleCellExperiment(so)
table(rownames(sce_fast)%in%rownames(sce))
assays(sce_fast) = assays(sce_fast)['logcounts']

# run singleR on Fast Louvain.Cluste annotation
pred_fast = SingleR(test = sce, ref = sce_fast, labels = sce_fast$Louvain.Cluster)
# add to single cell object
sce$cell_type_fast_Louvain_Cluster = toupper(as.character(pred_fast$pruned.labels)) 
# run singleR on Fast surface marker annotation
pred_fast2 <- SingleR(test = sce, ref = sce_fast, labels = sce_fast$surface_marker)
# add to single cell object
sce$cell_type_fast_surface_marker <- toupper(as.character(pred_fast2$pruned.labels)) 

```



# Integrating with protein abundance 

Gating stategy (Fast et al. 2021). See material and method for details

```{r gating}

## CD34
# Positive and negative to CD34
adt_cd34 = adt_matrix[which( adt_matrix[,'CD150'] > 5.5 & adt_matrix[,'CD48'] < 8),'CD34'] # binomial distribution

cut_30 = sort(adt_cd34,decreasing = F)[round(length(adt_cd34)/100*20)]
negativ = which(adt_cd34 < cut_30)
positiv = which(adt_cd34 >= cut_30)


gating_fast_LTHSC =  rownames(adt_matrix)[which( adt_matrix[,'CD34'] < cut_30 & adt_matrix[,'CD150'] > 5.5 & adt_matrix[,'CD48'] < 8)] # 753 (0.72%)
gating_fast_MPP1 = rownames(adt_matrix)[which( adt_matrix[,'CD34'] >= cut_30 & adt_matrix[,'CD150'] > 5.5 & adt_matrix[,'CD48'] < 8)] # 3.3
gating_fast_MPP2 =  rownames(adt_matrix)[which( adt_matrix[,'CD150'] > 5.5 & adt_matrix[,'CD48'] > 8)] # 10%
gating_fast_MPP34 =  rownames(adt_matrix)[which( adt_matrix[,'CD150'] < 5.5 & adt_matrix[,'CD48'] > 8)] # 43%
gating_fast_MPP0 =  rownames(adt_matrix)[which( adt_matrix[,'CD150'] < 5.5 & adt_matrix[,'CD48'] < 8)] # 10%


sce$gating_fast = 'no_gate'
sce$gating_fast[which(colnames(sce)%in%gating_fast_LTHSC)] = 'LTHSC'
sce$gating_fast[which(colnames(sce)%in%gating_fast_MPP1)] = 'STHSC'
sce$gating_fast[which(colnames(sce)%in%gating_fast_MPP2)] = 'MPP2'
sce$gating_fast[which(colnames(sce)%in%gating_fast_MPP34)] = 'MPP3'
sce$gating_fast[which(colnames(sce)%in%gating_fast_MPP0)] = 'MPP0'


# color accorinding to annotation
color_vector = sce$gating_fast
color_vector[which(color_vector=='MPP2')] = "#4363d8" 
color_vector[which(color_vector=='MPP3')] = "#3cb44b" 
color_vector[which(color_vector=='no_gate')] = "grey90" 
color_vector[which(color_vector=='LTHSC')] = "#e6194b"
color_vector[which(color_vector=='STHSC')] = "#f58231"
color_vector[which(color_vector=='MPP0')] = "#e5f531"

```


# Cluster filtering

Cluster 17, 18 and 19 were removed from the dataset prior to differential expression analysis (see material and method)

```{r filtering, echo=FALSE}
sce = sce[,-which(sce$final_clustering%in%c('17','18','19'))]
```

# Differential analysis

Pseudo-bulk analysis for each the cluster

```{r Differential_expression, echo=FALSE}
# Pseudo-replicate for pseudo-bulk analysis
sce$full_info_rev = paste0(sce$day,'_',sce$condition,'_',sce$batchNr)
# Rename clusters 0 to 9
sce$combined_clustering = paste('clust',sce$combined_clustering ,sep='')
sce$combined_clustering[which(sce$combined_clustering=='clust0.0')] = 'clust000'
sce$combined_clustering[which(sce$combined_clustering=='clust0.1')] = 'clust001'
sce$combined_clustering[which(sce$combined_clustering=='clust0.2')] = 'clust002'
sce$combined_clustering[which(sce$combined_clustering=='clust1')] = 'clust01'
sce$combined_clustering[which(sce$combined_clustering=='clust2')] = 'clust02'
sce$combined_clustering[which(sce$combined_clustering=='clust3')] = 'clust03'
sce$combined_clustering[which(sce$combined_clustering=='clust4')] = 'clust04'
sce$combined_clustering[which(sce$combined_clustering=='clust5')] = 'clust05'
sce$combined_clustering[which(sce$combined_clustering=='clust6')] = 'clust06'
sce$combined_clustering[which(sce$combined_clustering=='clust7')] = 'clust07'
sce$combined_clustering[which(sce$combined_clustering=='clust8')] = 'clust08'
sce$combined_clustering[which(sce$combined_clustering=='clust9')] = 'clust09'



splitCells = factor(paste0(make.names(sce$combined_clustering), ".", sce$full_info_rev))
num_cells <- table(splitCells)

# Filter: Exlude condition with less than 20 cell per samples
excludedCluster = names(num_cells[num_cells < 20]) 
sce.cell = sce[,-which(splitCells%in%excludedCluster)]
splitCells <- factor(paste0(make.names(sce.cell$combined_clustering), ".", sce.cell$full_info_rev))
# number of cells per samples and condition
num_cells <- table(splitCells)


## Sum of all cell from each sample
sum_by_cluster <- vapply(split(colnames(sce.cell), splitCells), function(x){ Matrix::rowSums(counts(sce.cell)[, x]) }, numeric(length = nrow(counts(sce.cell)))) 


# Create data frame containing information about the samples 
pheno <- data.frame(summed_sample=colnames(sum_by_cluster))
row.names(pheno) <- pheno$summed_sample
pheno$num_cells <- num_cells
pheno.comp = pheno
clu.day = unlist(lapply(pheno$summed_sample, function(x) strsplit(x,'_')[[1]][1]))
# cluster 
pheno.comp$clu = unlist(lapply(pheno$summed_sample, function(x) strsplit(x,'.',fixed = T)[[1]][1]))
# sample name
pheno.comp$sam = unlist(lapply(pheno$summed_sample, function(x) strsplit(x,'.',fixed = T)[[1]][2]))
# day 
pheno.comp$day = unlist(lapply(pheno.comp$sam, function(x) strsplit(x,'_',fixed=T)[[1]][1]))
# treatment
pheno.comp$tpo = unlist(lapply(pheno.comp$sam, function(x) strsplit(x,'_',fixed=T)[[1]][2]))
# fusion/no fusion
pheno.comp$fusion =unlist(lapply(pheno.comp$sam, function(x) strsplit(x,'_',fixed=T)[[1]][3]))
# full condition: day + treatment + fusion
pheno.comp$condition = paste(pheno.comp$day,pheno.comp$tpo,pheno.comp$fusion,sep='_')
# Add these info to final data frame
pheno$cluster <- factor(pheno.comp$clu)
pheno$replicate <- factor(pheno.comp$sam)
pheno$condition <- factor(pheno.comp$condition)

```

```{r DGEList, echo=FALSE}
## Create DGEList object 
# Create unique row names
row.names(sum_by_cluster) = uniquifyFeatureNames(rowData(sce)$GENEID, rowData(sce)$SYMBOL)
# combine info into dgelist
dge = DGEList(counts=sum_by_cluster,
             group=factor(paste(pheno$cluster, pheno$condition, sep=".")),
             samples=pheno,
             genes=as.data.frame(rowData(sce.cell)[, c('GENEID','SYMBOL','ENTREZID','uniqueSymbol','DESCRIPTION')]), 
             remove.zeros=FALSE)

# calculate normalization factors
dge = calcNormFactors(dge)

## Gene filtering
# Keep only the genes that are expressed in at least 2 samples
keep = rowSums(edgeR::cpm(dge) > 1) >= 2
# Subset, while updating total counts
dge = dge[keep, , keep.lib.sizes=FALSE] 
# Redo normalization (TMM)
dge = calcNormFactors(dge) 
# Clean-up 
row.names(dge) = dge$genes$GENEID 
# keep only most informative information
dge$genes = dge$genes[, c(1, 4, 3, 5)]
names(dge$genes) = c("GENEID", "SYMBOL", "ENTREZID", "DESCRIPTION")
# Visualize CPM after filtering
log2cpm.raw = edgeR::cpm(dge, log=TRUE, prior.count=8, normalized.lib.sizes=FALSE)
# and after TMM normalization
log2cpm = edgeR::cpm(dge, log=TRUE, prior.count=8, normalized.lib.sizes=TRUE) 


## design matrix 
moma = model.matrix(~ 0 + group, data=dge$samples) 
colnames(moma) = gsub("group", "", colnames(moma)) 
colnames(moma) = make.names(colnames(moma))

# Contrast matrix: cluster 0.02 can only be assessed for day2 (not enough replicate at day5)
contrasts.matrix <- makeContrasts(
  
  # cluster 000: 
  clust000.fusion.day2 = clust000.day2_TPO_FUSION - clust000.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust000.nofusion.day2 = clust000.day2_TPO_noFUSION - clust000.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust000.tpo.day2  = clust000.day2_TPO_FUSION - clust000.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust000.pbs.day2  = clust000.day2_PBS_FUSION - clust000.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # cluster 001: 
  clust001.fusion.day2 = clust001.day2_TPO_FUSION - clust001.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust001.nofusion.day2 = clust001.day2_TPO_noFUSION - clust001.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust001.tpo.day2  = clust001.day2_TPO_FUSION - clust001.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust001.pbs.day2  = clust001.day2_PBS_FUSION - clust001.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # cluster 002: 
  clust002.fusion.day2 = clust002.day2_TPO_FUSION - clust002.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust002.nofusion.day2 = clust002.day2_TPO_noFUSION - clust002.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust002.tpo.day2  = clust002.day2_TPO_FUSION - clust002.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust002.pbs.day2  = clust002.day2_PBS_FUSION - clust002.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)

  # clust01: 
  clust01.fusion.day2 = clust01.day2_TPO_FUSION - clust01.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust01.nofusion.day2 = clust01.day2_TPO_noFUSION - clust01.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust01.tpo.day2  = clust01.day2_TPO_FUSION - clust01.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust01.pbs.day2  = clust01.day2_PBS_FUSION - clust01.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # clust02: 
  clust02.fusion.day2 = clust02.day2_TPO_FUSION - clust02.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust02.nofusion.day2 = clust02.day2_TPO_noFUSION - clust02.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust02.tpo.day2  = clust02.day2_TPO_FUSION - clust02.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust02.pbs.day2  = clust02.day2_PBS_FUSION - clust02.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # clust03: 
  clust03.fusion.day2 = clust03.day2_TPO_FUSION - clust03.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust03.nofusion.day2 = clust03.day2_TPO_noFUSION - clust03.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust03.tpo.day2  = clust03.day2_TPO_FUSION - clust03.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust03.pbs.day2  = clust03.day2_PBS_FUSION - clust03.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # clust04: 
  clust04.fusion.day2 = clust04.day2_TPO_FUSION - clust04.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust04.nofusion.day2 = clust04.day2_TPO_noFUSION - clust04.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust04.tpo.day2  = clust04.day2_TPO_FUSION - clust04.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust04.pbs.day2  = clust04.day2_PBS_FUSION - clust04.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
    
  # clust05: 
  clust05.fusion.day2 = clust05.day2_TPO_FUSION - clust05.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust05.nofusion.day2 = clust05.day2_TPO_noFUSION - clust05.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust05.tpo.day2  = clust05.day2_TPO_FUSION - clust05.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust05.pbs.day2  = clust05.day2_PBS_FUSION - clust05.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
     
  # clust06: 
  clust06.fusion.day2 = clust06.day2_TPO_FUSION - clust06.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust06.nofusion.day2 = clust06.day2_TPO_noFUSION - clust06.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust06.tpo.day2  = clust06.day2_TPO_FUSION - clust06.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust06.pbs.day2  = clust06.day2_PBS_FUSION - clust06.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
     
  # clust07: 
  clust07.fusion.day2 = clust07.day2_TPO_FUSION - clust07.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust07.nofusion.day2 = clust07.day2_TPO_noFUSION - clust07.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust07.tpo.day2  = clust07.day2_TPO_FUSION - clust07.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust07.pbs.day2  = clust07.day2_PBS_FUSION - clust07.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
     
  # clust08: 
  clust08.fusion.day2 = clust08.day2_TPO_FUSION - clust08.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust08.nofusion.day2 = clust08.day2_TPO_noFUSION - clust08.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust08.tpo.day2  = clust08.day2_TPO_FUSION - clust08.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust08.pbs.day2  = clust08.day2_PBS_FUSION - clust08.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
     
  # clust09: 
  clust09.fusion.day2 = clust09.day2_TPO_FUSION - clust09.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust09.nofusion.day2 = clust09.day2_TPO_noFUSION - clust09.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust09.tpo.day2  = clust09.day2_TPO_FUSION - clust09.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust09.pbs.day2  = clust09.day2_PBS_FUSION - clust09.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
     
  # clust10: 
  clust10.fusion.day2 = clust10.day2_TPO_FUSION - clust10.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust10.nofusion.day2 = clust10.day2_TPO_noFUSION - clust10.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust10.tpo.day2  = clust10.day2_TPO_FUSION - clust10.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust10.pbs.day2  = clust10.day2_PBS_FUSION - clust10.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
     
  # clust11: 
  clust11.fusion.day2 = clust11.day2_TPO_FUSION - clust11.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust11.nofusion.day2 = clust11.day2_TPO_noFUSION - clust11.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust11.tpo.day2  = clust11.day2_TPO_FUSION - clust11.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust11.pbs.day2  = clust11.day2_PBS_FUSION - clust11.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # clust12: 
  clust12.fusion.day2 = clust12.day2_TPO_FUSION - clust12.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust12.nofusion.day2 = clust12.day2_TPO_noFUSION - clust12.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust12.tpo.day2  = clust12.day2_TPO_FUSION - clust12.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust12.pbs.day2  = clust12.day2_PBS_FUSION - clust12.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # clust13: 
  clust13.fusion.day2 = clust13.day2_TPO_FUSION - clust13.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust13.nofusion.day2 = clust13.day2_TPO_noFUSION - clust13.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust13.tpo.day2  = clust13.day2_TPO_FUSION - clust13.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust13.pbs.day2  = clust13.day2_PBS_FUSION - clust13.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # clust14: 
  clust14.fusion.day2 = clust14.day2_TPO_FUSION - clust14.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust14.nofusion.day2 = clust14.day2_TPO_noFUSION - clust14.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust14.tpo.day2  = clust14.day2_TPO_FUSION - clust14.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust14.pbs.day2  = clust14.day2_PBS_FUSION - clust14.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # clust15: 
  clust15.fusion.day2 = clust15.day2_TPO_FUSION - clust15.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust15.nofusion.day2 = clust15.day2_TPO_noFUSION - clust15.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust15.tpo.day2  = clust15.day2_TPO_FUSION - clust15.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust15.pbs.day2  = clust15.day2_PBS_FUSION - clust15.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # clust16: 
  clust16.fusion.day2 = clust16.day2_TPO_FUSION - clust16.day2_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust16.nofusion.day2 = clust16.day2_TPO_noFUSION - clust16.day2_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust16.tpo.day2  = clust16.day2_TPO_FUSION - clust16.day2_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust16.pbs.day2  = clust16.day2_PBS_FUSION - clust16.day2_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
    
  # cluster 000: 
  clust000.fusion.day5 = clust000.day5_TPO_FUSION - clust000.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust000.nofusion.day5 = clust000.day5_TPO_noFUSION - clust000.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust000.tpo.day5  = clust000.day5_TPO_FUSION - clust000.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust000.pbs.day5  = clust000.day5_PBS_FUSION - clust000.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # cluster 001: 
  clust001.fusion.day5 = clust001.day5_TPO_FUSION - clust001.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust001.nofusion.day5 = clust001.day5_TPO_noFUSION - clust001.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust001.tpo.day5  = clust001.day5_TPO_FUSION - clust001.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust001.pbs.day5  = clust001.day5_PBS_FUSION - clust001.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  

  # clust01: 
  clust01.fusion.day5 = clust01.day5_TPO_FUSION - clust01.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust01.nofusion.day5 = clust01.day5_TPO_noFUSION - clust01.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust01.tpo.day5  = clust01.day5_TPO_FUSION - clust01.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust01.pbs.day5  = clust01.day5_PBS_FUSION - clust01.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # clust02: 
  clust02.fusion.day5 = clust02.day5_TPO_FUSION - clust02.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust02.nofusion.day5 = clust02.day5_TPO_noFUSION - clust02.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust02.tpo.day5  = clust02.day5_TPO_FUSION - clust02.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust02.pbs.day5  = clust02.day5_PBS_FUSION - clust02.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # clust03: 
  clust03.fusion.day5 = clust03.day5_TPO_FUSION - clust03.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust03.nofusion.day5 = clust03.day5_TPO_noFUSION - clust03.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust03.tpo.day5  = clust03.day5_TPO_FUSION - clust03.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust03.pbs.day5  = clust03.day5_PBS_FUSION - clust03.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # clust04: 
  clust04.fusion.day5 = clust04.day5_TPO_FUSION - clust04.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust04.nofusion.day5 = clust04.day5_TPO_noFUSION - clust04.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust04.tpo.day5  = clust04.day5_TPO_FUSION - clust04.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust04.pbs.day5  = clust04.day5_PBS_FUSION - clust04.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
    
  # clust05: 
  clust05.fusion.day5 = clust05.day5_TPO_FUSION - clust05.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust05.nofusion.day5 = clust05.day5_TPO_noFUSION - clust05.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust05.tpo.day5  = clust05.day5_TPO_FUSION - clust05.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust05.pbs.day5  = clust05.day5_PBS_FUSION - clust05.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
     
  # clust06: 
  clust06.fusion.day5 = clust06.day5_TPO_FUSION - clust06.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust06.nofusion.day5 = clust06.day5_TPO_noFUSION - clust06.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust06.tpo.day5  = clust06.day5_TPO_FUSION - clust06.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust06.pbs.day5  = clust06.day5_PBS_FUSION - clust06.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
     
  # clust07: 
  clust07.fusion.day5 = clust07.day5_TPO_FUSION - clust07.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust07.nofusion.day5 = clust07.day5_TPO_noFUSION - clust07.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust07.tpo.day5  = clust07.day5_TPO_FUSION - clust07.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust07.pbs.day5  = clust07.day5_PBS_FUSION - clust07.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
     
  # clust08: 
  clust08.fusion.day5 = clust08.day5_TPO_FUSION - clust08.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust08.nofusion.day5 = clust08.day5_TPO_noFUSION - clust08.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust08.tpo.day5  = clust08.day5_TPO_FUSION - clust08.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust08.pbs.day5  = clust08.day5_PBS_FUSION - clust08.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
     
  # clust09: 
  clust09.fusion.day5 = clust09.day5_TPO_FUSION - clust09.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust09.nofusion.day5 = clust09.day5_TPO_noFUSION - clust09.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust09.tpo.day5  = clust09.day5_TPO_FUSION - clust09.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust09.pbs.day5  = clust09.day5_PBS_FUSION - clust09.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
     
  # clust10: 
  clust10.fusion.day5 = clust10.day5_TPO_FUSION - clust10.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust10.nofusion.day5 = clust10.day5_TPO_noFUSION - clust10.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust10.tpo.day5  = clust10.day5_TPO_FUSION - clust10.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust10.pbs.day5  = clust10.day5_PBS_FUSION - clust10.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
     
  # clust11: 
  clust11.fusion.day5 = clust11.day5_TPO_FUSION - clust11.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust11.nofusion.day5 = clust11.day5_TPO_noFUSION - clust11.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust11.tpo.day5  = clust11.day5_TPO_FUSION - clust11.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust11.pbs.day5  = clust11.day5_PBS_FUSION - clust11.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # clust12: 
  clust12.fusion.day5 = clust12.day5_TPO_FUSION - clust12.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust12.nofusion.day5 = clust12.day5_TPO_noFUSION - clust12.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust12.tpo.day5  = clust12.day5_TPO_FUSION - clust12.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust12.pbs.day5  = clust12.day5_PBS_FUSION - clust12.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # clust13: 
  clust13.fusion.day5 = clust13.day5_TPO_FUSION - clust13.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust13.nofusion.day5 = clust13.day5_TPO_noFUSION - clust13.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust13.tpo.day5  = clust13.day5_TPO_FUSION - clust13.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust13.pbs.day5  = clust13.day5_PBS_FUSION - clust13.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # clust14: 
  clust14.fusion.day5 = clust14.day5_TPO_FUSION - clust14.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust14.nofusion.day5 = clust14.day5_TPO_noFUSION - clust14.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust14.tpo.day5  = clust14.day5_TPO_FUSION - clust14.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust14.pbs.day5  = clust14.day5_PBS_FUSION - clust14.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # clust15: 
  clust15.fusion.day5 = clust15.day5_TPO_FUSION - clust15.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust15.nofusion.day5 = clust15.day5_TPO_noFUSION - clust15.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust15.tpo.day5  = clust15.day5_TPO_FUSION - clust15.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust15.pbs.day5  = clust15.day5_PBS_FUSION - clust15.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
  # clust16: 
  clust16.fusion.day5 = clust16.day5_TPO_FUSION - clust16.day5_PBS_FUSION, # TPO vs. PBS when fusion (overexpression in TPO)
  clust16.nofusion.day5 = clust16.day5_TPO_noFUSION - clust16.day5_PBS_noFUSION, # TPO vs. PBS when no fusion (overexpression in TPO)
  clust16.tpo.day5  = clust16.day5_TPO_FUSION - clust16.day5_TPO_noFUSION, # fusion vs no fusion when under TPO (overexpression in fusion)
  clust16.pbs.day5  = clust16.day5_PBS_FUSION - clust16.day5_PBS_noFUSION, # fusion vs no fusion when under PBS(overexpression in fusion)
  
 levels=moma
)



## Differential expression test per cluster

# save the results of DE in list
fits <- list()

# Loop across the contrast
for (i in 1:ncol(contrasts.matrix)) {
  
  # Name of the constrast
  contr.name <- colnames(contrasts.matrix)[i]
  
  # Print info
  message("   Working on ... ", contr.name)
  
  # Subset samples
  dge.temp <- dge[, dge$samples$group %in% names(which(contrasts.matrix[,i] != 0))]
  
  ## Subset and renormalize
  keep <- rowSums(edgeR::cpm(dge.temp) > 1) >= 2  
  dge.temp <- dge.temp[keep, , keep.lib.sizes=FALSE] 
  
  ## Add a filtering step on the min number of cells with expression!
  ## keep only the cell belonging to the contrast we are making
  sce.temp <- 
    sce[
      rowData(sce)$GENEID %in% dge.temp$genes$GENEID, 
      paste0(sce$combined_clustering, ".",sce$day,'_', sce$TPO,'_',sce$Fusion) %in% names(which(contrasts.matrix[,i] != 0)) ]
  
  ## Should be detected in at least 5% of the cells
  keep <- rowSums(counts(sce.temp) > 0 ) > dim(sce.temp)[2]/100*5
  dge.temp <- dge.temp[keep, , keep.lib.sizes=FALSE] 
  
  # print(dim(dge.temp)) # total 5568 genes for MF9_d60.clust10vs1
  
  ## Redo TMM normalization
  dge.temp <- calcNormFactors(dge.temp) 
  ## After TMM normalization
  log2cpm.temp <- edgeR::cpm(dge.temp, log=TRUE, prior.count=8, normalized.lib.sizes=TRUE)


  ## design matrix 
  moma.temp <- model.matrix(~ 0 + group, data=dge.temp$samples) 
  colnames(moma.temp) <- gsub("group", "", colnames(moma.temp)) 
  colnames(moma.temp) <- make.names(colnames(moma.temp))
  colnames(moma.temp)
  
  ## estimate dispersion (TMM norm object)
  dge.temp <- estimateDisp(dge.temp, moma.temp)

  fit1 <- glmFit(dge.temp, moma.temp, prior.count=8) 
  lrt <- glmLRT(fit1, contrast=contrasts.matrix[names(which(contrasts.matrix[,i] != 0)), i])


  ## add results to list
  fits[[contr.name]] <- lrt
  rm(lrt)
  
  ## Save filtered DGEList and other info useful for later 
  fits[[contr.name]]$dge <- dge.temp
  fits[[contr.name]]$moma <- moma.temp

}


```



# Figure

```{r}
# final figure
myTheme =  theme_bw(base_size = 20) + 
  theme(panel.grid.major = element_line(colour = "lightgrey", linewidth = 0.1))  +
  theme(axis.text=element_text(size=20))

sce$cell_type_fast_Louvain_Cluster = str_replace(sce$cell_type_fast_Louvain_Cluster ,'LSK-','')

a = plotReducedDim(sce, dimred="TSNE_MNN_Sample_k50", colour_by="final_clustering" ) + myTheme + xlab('TSNE.1') + ylab('TSNE.2') + coord_fixed( ) 

b = plotReducedDim(sce, dimred="TSNE_MNN_Sample_k50", colour_by="cell_type_fast_Louvain_Cluster" )+  scale_color_manual(name = "",values=c('PRIMITIVE'='#296DA0','PRIMED'='#ADB55F','METABOLISM'='#259461','CELL-CYCLE'='#7E5146','INTERFERON CELL-CYCLE'='#C971A1','INTERFERON'='#795291','ACUTE-ACTIVATION'='#C32B2C','MYELOID'='#31B6BF'))  + myTheme + xlab('TSNE.1') + ylab('TSNE.2' ) + coord_fixed( ) 

c = plotReducedDim(sce, dimred="TSNE_MNN_Sample_k50", colour_by="phases" ) + myTheme + xlab('TSNE.1') + ylab('TSNE.2') + coord_fixed( ) 


colors = c (LTHSC="#e6194b",MPP0= "grey90",STHSC="grey90",MPP2="grey90",MPP3="grey90" ,no_gate='grey90')
LTHSC = plotReducedDim(sce[,which(sce$gating_fast=='LTHSC')], dimred="TSNE_MNN_Sample_k50", colour_by="gating_fast", size_by='gating_fast',shape_by = 'gating_fast') +  myTheme + coord_fixed( ) +
  scale_color_manual(name = "gating_fast",values=colors) + scale_size_manual(name = "gating_fast",values=c (LTHSC=1,MPP0= 0.01,STHSC=0.1,MPP2=0.01,MPP3=0.01 ,no_gate=0.01)) +  scale_shape_manual(name = "gating_fast",values=c (LTHSC=16,MPP0= 16,STHSC=16,MPP2=16,MPP3=16 ,no_gate=16)) + geom_density_2d(color='grey80') + xlab('TSNE.1') + ylab('TSNE.2') 


colors = c (LTHSC="grey90",MPP0 ="#e5f531",STHSC="grey90",MPP2="grey90",MPP3="grey90" ,no_gate='grey90')
MPP0 =plotReducedDim(sce[,which(sce$gating_fast=='MPP0')], dimred="TSNE_MNN_Sample_k50", colour_by="gating_fast", size_by='gating_fast',shape_by = 'gating_fast') +  myTheme + coord_fixed( ) +
  scale_color_manual(name = "gating_fast",values=colors) + scale_size_manual(name = "gating_fast",values=c (LTHSC=0.01,MPP0= 1,STHSC=0.1,MPP2=0.01,MPP3=0.01 ,no_gate=0.01)) +  scale_shape_manual(name = "gating_fast",values=c (LTHSC=16,MPP0= 16,STHSC=16,MPP2=16,MPP3=16 ,no_gate=16)) + geom_density_2d(color='grey80') + xlab('TSNE.1') + ylab('TSNE.2')

colors = c (LTHSC="grey90",MPP0= "grey90",STHSC="#f58231",MPP2="grey90",MPP3="grey90" ,no_gate='grey90')
STHSC = plotReducedDim(sce[,which(sce$gating_fast=='STHSC')], dimred="TSNE_MNN_Sample_k50", colour_by="gating_fast", size_by='gating_fast',shape_by = 'gating_fast') +  myTheme + coord_fixed( ) +
  scale_color_manual(name = "gating_fast",values=colors) + scale_size_manual(name = "gating_fast",values=c (LTHSC=0.01,MPP0= 0.01,STHSC=1,MPP2=0.01,MPP3=0.01 ,no_gate=0.01)) +  scale_shape_manual(name = "gating_fast",values=c (LTHSC=16,MPP0= 16,STHSC=16,MPP2=16,MPP3=16 ,no_gate=16)) + geom_density_2d(color='grey80') + xlab('TSNE.1') + ylab('TSNE.2')

colors = c (LTHSC="grey90",MPP0= "grey90",STHSC="grey90",MPP2="#4363d8",MPP3="grey90" ,no_gate='grey90')
MPP2 = plotReducedDim(sce[,which(sce$gating_fast=='MPP2')], dimred="TSNE_MNN_Sample_k50", colour_by="gating_fast", size_by='gating_fast',shape_by = 'gating_fast') +  myTheme + coord_fixed( ) +
  scale_color_manual(name = "gating_fast",values=colors) + scale_size_manual(name = "gating_fast",values=c (LTHSC=0.01,MPP0= 0.01,STHSC=0.01,MPP2=1,MPP3=0.01 ,no_gate=0.01))  +  scale_shape_manual(name = "gating_fast",values=c (LTHSC=16,MPP0= 16,STHSC=16,MPP2=16,MPP3=16 ,no_gate=16)) + geom_density_2d(color='grey80') + xlab('TSNE.1') + ylab('TSNE.2')


colors = c (LTHSC="grey90",MPP0= "grey90",STHSC="grey90",MPP2="grey90",MPP3="#3cb44b" ,no_gate='grey90')
MPP3 = plotReducedDim(sce[,which(sce$gating_fast=='MPP3')], dimred="TSNE_MNN_Sample_k50", colour_by="gating_fast", size_by='gating_fast',shape_by = 'gating_fast') +  myTheme + coord_fixed( ) +
  scale_color_manual(name = "gating_fast",values=colors) + scale_size_manual(name = "gating_fast",values=c (LTHSC=0.01,MPP0= 0.01,STHSC=0.01,MPP2=0.01,MPP3=1 ,no_gate=0.01))  +  scale_shape_manual(name = "gating_fast",values=c (LTHSC=16,MPP0= 16,STHSC=16,MPP2=16,MPP3=16 ,no_gate=16)) + geom_density_2d(color='grey80') + xlab('TSNE.1') + ylab('TSNE.2')

```
